### 2016-10-24 Manuscript/SW_early_life_telomeres.Rmd
More edits

5. Our comprehensive, long-term data show that in natural populations telomere dynamics vary enormously over both space and time, and that this variation cannot be attributed entirely to measurment error. Ascertaining what explains this variation will be a difficult but worthwhile task, and will require combining long-term ecological and quantitative genetic data.

@@

#####Early-life telomere dynamics and survival
Early-life RTL was not related to survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~cenTL+Agemonths+(1|PlateID)+(1|TerritoryID)+(1|FieldPeriodID),data=juv14,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glmer(SurvivedNext~cenTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv14,family = 'binomial'))$coefficients[2,4])`), with no interaction effect between RTL and juvenile age (estimate = `r round2(summary(glmer(SurvivedNext~RTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv14,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glmer(SurvivedNext~RTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv14,family = 'binomial'))$coefficients[4,4])`), suggesting that there were no age-specific effects of juvenile RTL on survival to adulthood. To test whether cohort-level effects on RTL influenced survival we calculated cohort-level RTL (i.e. the median RTL for each cohort) and added this term to a GLMM, and found that cohort-level RTL was not related to survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~cohortTL+RTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv14,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glmer(SurvivedNext~cohortTL+RTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv14,family = 'binomial'))$coefficients[2,4])`).

@@

temp <- subset(dd3,BirdID %in% juv$BirdID)
nrow(temp)

>>>>

5. Our comprehensive, long-term data show that in natural populations telomere dynamics vary enormously over both space and time, and that this variation cannot be attributed entirely to measurment error. Ascertaining what explains this variation will require combining long-term ecological and quantitative genetic data.

@@

#####Early-life telomere dynamics and survival
Early-life RTL was not related to survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~cenTL+Agemonths+(1|PlateID)+(1|TerritoryID)+(1|LayYear),data=juv14,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glmer(SurvivedNext~cenTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv14,family = 'binomial'))$coefficients[2,4])`), with no interaction effect between RTL and juvenile age (estimate = `r round2(summary(glmer(SurvivedNext~RTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv14,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glmer(SurvivedNext~RTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv14,family = 'binomial'))$coefficients[4,4])`), suggesting that there were no age-specific effects of juvenile RTL on survival to adulthood. To test whether cohort-level effects on RTL influenced survival we calculated cohort-level RTL (i.e. the median RTL for each cohort) and added this term to a GLMM, and found that cohort-level RTL was not related to survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~cohortTL+RTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv14,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glmer(SurvivedNext~cohortTL+RTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv14,family = 'binomial'))$coefficients[2,4])`).

@@

m1 <- lmer(RTL~LogAge+(1|BirdID),data=dd)
m2 <- lmer(RTL~LogAge+(1|BirdID) + (1|LayYear),data=dd)
m3 <- lmer(RTL~LogAge+(1|BirdID) + (LogAge|LayYear),data=dd)

anova(m1,m2,m3)

dd14 <- subset(dd,DeathYear<2015)
dd14$RowID <- c(1:nrow(dd14))

summary(glmer(SurvivedNext~RTL * LogAge + Tarsus + Sex + (1|BirdID)+(1|LayYear),data=dd14,family = 'binomial'))

ggplot(dd14,aes(x = Agemonths, col = factor(SurvivedNext),y = RTL))+
geom_smooth(method = 'lm',formula = y~log(x))

### 2016-10-24 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd
More edits

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'

@@

Fig3B <- ggplot(temp,
       aes(x = Tarsus,
           y = RTL,
           col = Age2)) +
  ylab('RTL') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'top') +
  stat_smooth(method = 'lm')
  annotate('text',x = 17,y = 2.1,label = 'B')

temp$InsectF <- ifelse(temp$Insect>4,'High','Low')
Fig3C <- ggplot(temp,
       aes(
           y = RTL,
           x=Agemonths,
           col = InsectF)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'top') +
  geom_point()+
   stat_smooth(method = 'lm',formula = y~log(x))
  # scale_colour_manual(values = c('grey','darkorange','darkgreen'))+
  # scale_fill_manual(values = c('grey','darkorange','darkgreen'))+
  annotate('text',x = 3,y = 2.1,label = 'C')

@@

anova(lm(RTL~Sex+Agemonths+factor(CatchYear)+Sex*factor(LayYear),data=dd))

ggplot(dd,aes(x = factor(LayYear),y = RTL,col = Sex))+
geom_boxplot(notch=F)

temp <- subset(dd,Age<15)
reps <- tapply(temp$RTL,temp$BirdID,length)
keep <- names(reps[reps>2])
temp <- subset(temp,BirdID %in% keep)
 temp <- subset(temp,!duplicated(paste0(BirdID,Agemonths)))

ggplot(temp,aes(x = CatchYear,y = RTL,group = BirdID,col=factor(BirdID)))+
  geom_point(alpha = 1,col='grey')+
geom_line(col = 'darkgrey',alpha = 0.2)+
  xlab('Age (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none')+
  #scale_colour_brewer(palette = 'Set1')+
  annotate('text',x = 1997,y = 2,label = 'B')

ggplot(ddL,aes(x = DeltaRTL,y = RTL,col = Group))+
geom_point()+
stat_smooth(method = 'lm')

```{r Figure X}

temp <- dd
#temp$SurvivedNext <- factor(temp$SurvivedNext)
temp$SN2 <- with(temp,paste(Sex,SurvivedNext))

FigX <- ggplot(temp,aes(x = LayYear,y = RTL,col = SN2))+
  ylab('RTL') + 
  xlab('Cohort') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12,angle = 90), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'top')+
  geom_point()+
  stat_smooth()

FigX

FigX <- ggplot(subset(temp,Agemonths),aes(x = SN2,y = RTL))+
  ylab('RTL') + 
  xlab('Cohort') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12,angle = 90), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'top')+
  geom_boxplot()

FigX

temp <- dd
temp$LogAge <- log(temp$Agemonths)

m1 <- lm(RTL~LogAge,data=temp)
m2 <- lmer(RTL~LogAge+(1|CatchYear),data=temp)
m3 <- lmer(RTL~LogAge+(LogAge|CatchYear),data=temp)
m4 <- lmer(RTL~LogAge+(1|LayYear),data=temp)
m5 <- lmer(RTL~LogAge+(LogAge|LayYear),data=temp)

m1 <- lmer(RTL~(Agemonths)+(1|BirdID)+(1|LayYear),data=dd)
m2 <- lmer(RTL~Ageclass+(1|BirdID)+(1|LayYear),data=dd)
m3 <- lmer(RTL~LogAge+(1|BirdID)+(1|LayYear),data=dd)

anova(m1,m2,m3)
summary(m2)
summary(m3)

dd3$LogAge <- log(dd3$Agemonths)
m1 <- lmer(RTL1~RTL+(1|LayYear),data=dd3)
summary(m1)

>>>>

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.LogAge'] <- 'Log age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.CatchMonth'] <- 'Month'
nams[nams == 'z.Density'] <- 'Density'

@@

Fig3B <- ggplot(temp,
       aes(x = Agemonths,
           y = RTL,
           col = Sex)) +
  ylab('RTL') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'top') +
  stat_smooth(method = 'lm',formula = y~log(x))
  annotate('text',x = 17,y = 2.1,label = 'B')

Fig3C <- ggplot(temp,
       aes(
           y = RTL,
           x=Insect,
           col = Sex)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'top') +
  geom_point()+
   stat_smooth(method = 'lm',formula = y~log(x))
  # scale_colour_manual(values = c('grey','darkorange','darkgreen'))+
  # scale_fill_manual(values = c('grey','darkorange','darkgreen'))+
  annotate('text',x = 3,y = 2.1,label = 'C')

@@

plot(tapply(dd3$DeltaRTLF,dd3$TimeDiff,mean))
